---
title: "Analyses from the behavioural experiment"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
          message = FALSE,
         warning = FALSE,
         fig.width = 8, fig.height = 8)
library(ggplot2)
library(ggpubr)
library(lme4)
library(lmerTest)
library(effects)
library("plyr")
library(aods3) #to check goodness of fit - function gof
library(car) #to get Anova-like tables

```

```{r eval=FALSE, include=FALSE}
rm(list=ls())
Dane<-read.csv2("Data.csv")
```

```{r eval=FALSE, include=FALSE}
# variable name	- description
# eksp_population	- population names and replicates were encoded for the purpose of the experiment, for blinding the experimenters
# seria	- series; plates from one series were subject to observation in the same day
# first_contact	- time to first contact (s)
# contact	- occurrence of contact with any of the females (0-no, 1-yes)
# contact_duration -	duration of contact with the females (s)
# offspring	- offspring presence on the next day after observation (0-no, 1-yes)
# tail_chasing	- duration of the tail-chasing behaviour (s)
# population	- population name
# generation	- generation in experimental evolution from which the analysed population was taken from a frozen stock (anc - ancestral, 97 - 97th)

```

```{r echo=FALSE}
Dane$contact<-as.factor(Dane$contact)
Dane$offspring<-as.factor(Dane$offspring)
```

# I. Analyses assessing the differences among populations in the following traits:


(1)	Time to first contact 
```{r}
model1<-lm(first_contact~population, data=Dane)
summary(model1)
Anova(model1, type="III")

#zexcluding cases where contact did not occur
Dane$contact<-as.character(Dane$contact)
Dane.sel<-subset(Dane, contact=="1")

model1a<-lm(first_contact~population, data=Dane.sel)
summary(model1a)
Anova(model1a, type="III")

```


(2)	Occurrence (or not) of at least one contact with any of the females 
```{r}
model2<-glm(contact~population, data=Dane, family="binomial")
summary(model2)
Anova(model2, type="III")
plot(model2)
gof(model2) #checking goodness of fit
```

(3)	Total time spent in contact with females 
```{r}
model3<-lm(contact_duration~population, data=Dane)
summary(model3)
Anova(model3, type="III")
plot(model3)
```
 
(4)	Copulation success (presence of offspring the day after observation)
```{r}
model4<-glm(offspring~population, data=Dane, family="binomial")
summary(model4)
Anova(model4, type="III")
plot(model4)
gof(model4)
```

(5)	Time spent by the male on chasing its tail
```{r}
model5<-lm(tail_chasing~population, data=Dane)
summary(model5)
Anova(model5, type="III")
plot(model5)

```

#II. Analyses of relationships between traits, with population included as an additional fixed factor: 

(6)	The effect of  the time spent by the male on chasing its tail on the total duration of contact
```{r}
model6<-lm(contact_duration~tail_chasing+population, data=Dane)
summary(model6)
plot(Dane$tail_chasing,Dane$contact_duration)
abline(model6)

Anova(model6, type="III")
res<-resid(model6)
hist(res)

model6.int<-lm(contact_duration~tail_chasing+population+tail_chasing*population, data=Dane)
summary(model6.int)
#interaction is not significant

anova(model6.int, model6)
#checking if the model with interaction (model6.int) is not better than the model without (model6) - it is not

```





(7)	The effect of the time spent by the male on chasing its tail on copulation success (presence of offspring)
```{r}
model7<-glm(offspring~tail_chasing+population,  family="binomial", data=Dane)
summary(model7)
Anova(model7, type="III")
library(aods3)
gof(model7)

model7.int<-glm(offspring~tail_chasing+population+tail_chasing*population,  family="binomial", data=Dane)
summary(model7.int)
#interaction is not significant

anova(model7.int, model7, test="Chi")
```

(8)	The effect of the total duration of contact on copulation success
```{r}

model8<-glm(offspring~contact_duration+population,  family="binomial", data=Dane)
summary(model8)

Anova(model8, type="III")
plot(effect("contact_duration",model8))
gof(model8)
ggplot(Dane, aes(x=contact_duration, y=offspring))+
  geom_point()


model8.int<-glm(offspring~contact_duration+population+contact_duration*population,  family="binomial", data=Dane)
summary(model8.int)
Anova(model8.int, type="III")
#interaction is not significant

anova(model8.int, model8, test="Chi")
```


